Trade-offs between sustainable development goals in systems of cities

### 2022-04-04 paper.tex
Update on Overleaf.

%\author{}
\author[label1]{Juste Raimbault\corref{cor1}}
\author[label2]{Denise Pumain}

\address[label1]{CASA, University College London}
\address[label2]{G{\'e}ographie-cit{\'e}s, Universit{\'e} Paris 1}

\cortext[cor1]{Corresponding author: juste.raimbault@polytechnique.edu}

\begin{abstract}
    Sustainable Development Goals are intrinsically competing, but their embedding into urban systems furthermore emphasises such compromises, due to spatial complexity, the non-optimal nature of such systems, and multi-objective aspects of their agents, among other reasons. We propose in this paper to use a stylised simulation model for systems of cities, focused on innovation diffusion and population dynamics, to show how trade-offs may operate at such a scale. We proceed in particular to a bi-objective optimisation of emissions and innovation utilities, and show that no single urban optimum exists, but a diversity of regimes forming a compromise between the two objectives.
\end{abstract}

@@

Systems of cities are an almost universal form of spatial organization of societies to inhabit the earth space. This organization has proven to be very sustainable for several millennia, and has continuously adapted to the political, economic and technological innovations of societies over the centuries \citep{pumain2020theories} and perhaps now the largest may have become ``global rulers'' \citep{glaeser2020urban}. However, for the last few decades it seems to be threatened by excessive emissions and overexploitation of the planet's energy and material resources \citep{nijkamp2014sustainable,kourtit2020global}. While demographic and economic growth have supported and accelerated the development of urbanization and the proliferation of systems of cities for two centuries, they are now accused of being the cause of climate change problems and the rise of social inequalities \citep{davis2006planet,glaeser2009inequality}. It seems that excessive mobility associated with the fragmentation of the value chains of globalized production has progressively decoupled urban expansion from the proper management of planetary resources \citep{rozenblat2018urban}. It is unlikely that the technological solutions that are provided under the label of smart cities could easily solve these problems \citep{caragliu2011smart,kourtit2020global}.

A good knowledge of the dynamics of the systems of cities is useful to prepare possible interventions in urban systems because of their complexity \citep{reggiani2021reflections}. It implies a very important collection of information on the processes of urban dynamics at all geographical scales. The variables to be considered are considerably more numerous and the knowledge of their interactions is much less easy to build than those, already fantastically colossal, which were necessary to ascertain the evolutionary direction of the climate. Alongside the many localised investigations, we believe that abstract modelling, which distances itself from the diversity of cities, their cultures and subjective sensitivities, can help to discern some of the possible ways of managing urban dynamics \citep{pumain2017urban}.

Systems of cities are characterized by co-evolutionary processes with non-linear dynamics far from equilibrium, which makes forecasting attempts particularly difficult \citep{raimbault2020unveiling}. By integrating stylized facts from a large number of observations into simulation models, it is possible to list the most probable paths of their dynamics and to support reflection on possible evolution, without proposing a priori a horizon that would be more desirable than another. Indeed, we have often found that the diversity of cities, in size and function, is an important component of the adaptive dynamics of systems of cities \citep{pumain2021co}. The search for an optimum that would value adaptation to a standard or to the situation of the moment would necessarily be doomed to failure. Moreover, a specific attention should be paid to temporal scales: a recent attempt at modeling urban growth with empirical observations of daily mobility rediscovers that not only ``strong social interactions but also long-term memory effects'' are major principles for capturing urban dynamics \citep{xu2021emergence}.

The concept of urban optima, in the sense of optimising certain dimensions of urban systems, has been considered from diverse perspectives. It is often conceived within the economic paradigm of equilibrium \citep{glaeser2008cities}. In most cases, there does not seem to be clear patterns, neither empirical nor theoretical, of possible simple optimisations of single objectives by urban systems. Some results in urban economics regarding an optimal city size requires to consider a city in a closed system, which is unreasonable from a realistic perspectives \citep{singell1974optimum}. Studies of an optimal urban population density are restricted to economic criteria of wage and productivity \citep{su2017density}. The sustainability of urban forms for CO2 emissions requires considering complex indicators of urban form \citep{le2012urban}. Similarly, no universal rule seems to exist for the scaling of emissions with city size \citep{gudipudi2019urban}. In terms of pollution, empirical results across different urban systems suggest no fixed relationship between city size and emission of pollutants \citep{han2016optimum}. Altogether, this converges with the idea of multiple agents optimising multiple dimensions at different scales \citep{pumain2008socio}, and therefore no empirical support for simple ``urban optima''.

@@

The simulated urban system is composed by cities, characterised at each time step by (i) their population; (ii) their location in the geographical space; (iii) adoption rates by their populations for different innovations (urban genome).

@@

>>>>

\author{}
%\author[label1]{Juste Raimbault\corref{cor1}}
%\author[label2]{Denise Pumain}

%\address[label1]{CASA, University College London}
%\address[label2]{G{\'e}ographie-cit{\'e}s, Universit{\'e} Paris 1}

%\cortext[cor1]{Corresponding author: juste.raimbault@polytechnique.edu}

\begin{abstract}

    Sustainable Development Goals are intrinsically competing, and their embedding into urban systems furthermore emphasises such compromises. When observed at the scale of systems of cities, such concern is considered as a series of innovations that challenges the adaptive capacity of urban systems. The spatial complexity, the non-optimal nature of such systems, and the multi-objective aspects of their agents, are among the reasons that raise difficulties when trying to adjust local policies through promoting innovation in order to satisfy at least a couple of SDGs simultaneously. As we lack enough empirical evidence, we propose in this paper to use a stylised simulation model for systems of cities, focused on innovation diffusion and population dynamics, to show how trade-offs may operate at such a scale. We proceed in particular to a bi-objective optimisation of emissions and innovation utilities, and show that no single urban optimum exists, but a diversity of regimes forming a compromise between the two objectives.
\end{abstract}

@@

The 17 Sustainable Development Goals that were defined by the United Nations in 2015 \citep{nations2015sustainable} are challenging the adaptation capability of many cities in the world. As they host a growing proportion of the world population and are responsible for a large part of greenhouse gas emissions (Christen 2014), cities are at the core of environmental transition policies \citep{romero2011cities}. Indeed, cities are explicitly mentioned in SDG 11 \citep{nations2015sustainable}: ``Make cities and human settlements inclusive, safe, resilient, and sustainable''. This citation demonstrates that many other SDGs involving a diversity of urban issues and stakeholders are already embedded in that specific goal \citep{vaidya2020sdg}.

In the ongoing adaptive process of the environmental transition, cities are not isolated. They have been engaged for long in a co-opetition process for attracting population and economic activities that have created many interdependencies in their evolution leading to interpret cities as “systems within systems of cities” (Berry, 1964). These systems of cities are an almost universal form of spatial organization of societies to inhabit the earth space. This organization has proven to be very sustainable for several millennia, because it offered a way to continuously adapt the political, economic and technological innovations that societies imagined over the centuries \citep{pumain2020theories}. That remarkable ability to resilience through self-renewal has expanded all over the world up to the point that the largest metropolises may have now become ``global rulers'' \citep{glaeser2020urban}. However, for the last few decades urban systems seem to be threatened by excessive emissions and overexploitation of the planet's energy and material resources \citep{nijkamp2014sustainable,kourtit2020global}. While demographic and economic growth have supported and accelerated the development of urbanization and the proliferation of systems of cities for two centuries, they are now accused of being the cause of climate change problems and the rise of social inequalities \citep{davis2006planet,glaeser2009inequality}. It seems that the excessive mobility associated with the fragmentation of the value chains of globalized production has progressively decoupled urban expansion from the proper management of planetary resources \citep{rozenblat2018urban}. It is unlikely that the technological solutions that are provided under the label of “smart cities” could easily solve these problems \citep{caragliu2011smart,kourtit2020global}, especially because of the huge diversity of cities internal layouts and their already established networks operating internal and external interactions \citep{}(Caruso et al. 2022).

Regarding the adaptation process to SDGs, recent research follows two main lines, involving theoretical and empirical investigations.  On the theoretical side, the last three decades have produced a huge literature, mainly oriented toward assessing the relationship between gas emissions and GDP over time. The environmental Kuznets curve (EKC) hypothesis proposes that there is an inverted U-shape relation between environmental degradation and income per capita \citep{dinda2004environmental, stern2004rise}. This has been taken to imply that economic growth will eventually redress the environmental impacts observed from the early stages of economic development. However, difficulties linked with conceptual definitions and indicators measurements, although helping to clarify the issue, have not led to a complete consensus about the robustness of that hypothesis \citep{harbaugh2002reexamining}. Moreover, most tests regarding that hypothesis have been made at a macrogeographic level and compare national statistical evolutions. The observations made at city level are more recent and fragmented. Empirical investigations about SDGs in cities still lack of well-established and comparable sources of information. Thus they follow a diversity of directions: either identifying relevant indicators for benchmarking and monitoring the improvement process \citep{giles2020achieving}, or measuring their variations according to city size \citep{laituri2021sdg} and according to specific practices and specific national contexts (Mudau et al., 2020 for inclusivity South Africa, Mohan et al. 2021 for road safety in India, Koch and Krellenberg 2018 for Germany, Zhu et al 2022 for the relationship between carbon emissions and polycentric structures in China).

Alongside the many localised investigations, we believe that abstract modelling, which distances itself from the diversity of cities, their cultures and subjective sensitivities, can help as a first step to discern some of the possible ways of managing urban dynamics \citep{pumain2017urban}. A good knowledge of the dynamics of the systems of cities is required to prepare possible interventions in urban systems because of their intrinsic complexity\citep{reggiani2021reflections}. Systems of cities are characterized by co-evolutionary processes with non-linear dynamics far from equilibrium, which makes forecasting attempts particularly difficult \citep{raimbault2020unveiling}. By integrating stylized facts from a large number of observations into simulation models, it is possible to list the most probable paths of their dynamics and to support reflection on possible evolution, without proposing a priori a horizon that would be more desirable than another. Indeed, we have often found that the diversity of cities, in size and function, is an important component of the adaptive dynamics of systems of cities \citep{pumain2021co}. The search for an optimum that would value adaptation to a standard or to the situation of the moment would necessarily be doomed to failure. Moreover, a specific attention should be paid to temporal scales: a recent attempt at modeling urban growth with empirical observations of daily mobility rediscovers that not only ``strong social interactions but also long-term memory effects'' are major principles for capturing urban dynamics \citep{xu2021emergence}.

The concept of urban optima, in the sense of optimising certain dimensions of urban systems, has been considered from diverse perspectives. It is often conceived within the economic paradigm of equilibrium \citep{glaeser2008cities}. In most cases, there does not seem to be clear patterns, neither empirical nor theoretical, of possible simple optimisation of single objectives by urban systems. Some results in urban economics regarding an optimal city size requires to consider a city in a closed system, which is unreasonable from a realistic perspectives \citep{singell1974optimum}. Studies of an optimal urban population density are restricted to economic criteria of wage and productivity \citep{su2017density}. The sustainability of urban forms for CO2 emissions requires considering complex indicators of urban form \citep{le2012urban}. Similarly, no universal rule seems to exist for the scaling of emissions with city size \citep{gudipudi2019urban}. In terms of pollution, empirical results across different urban systems suggest no fixed relationship between city size and emission of pollutants \citep{han2016optimum}. Altogether, this converges with the idea of multiple agents optimising multiple dimensions at different scales \citep{pumain2008socio}, and therefore no empirical support for simple ``urban optima''.

@@

The simulated urban system is composed by cities, which location in the geographical space and number are fixed in time. They are characterised at each time step by their population $P_i (t)$ and by adoption rates by their populations for different innovations (urban genome $\delta_{i,c} (t)$).

@@

% additional dimensions? wage inequalities \cite{shutters2022urbanization}

% on kuznet debate: invest in what is working now \cite{brisbois2022climate}
